Chatbots and messaging platforms in the classroom: an analysis from the teacher's perspective

### 2021-07-05 edubots_surveys.Rnw
Update edubots_surveys.Rnw

Given the potential advantages that chatbots could bring to the classroom, as a previous step, we created two surveys for bachelor and master degree students at the University of Granada (Spain), which were answered by more than 250 students. The obtained results were analysed with the object of identifying students' actual needs, in \cite{MoraChatbots2021}. The survey asked about their preferences on tools/applications for chatting and messaging, how do they use them in the educational context, who do they like to be with in the class messaging groups, and what do they expect bots and chatbots to do to aid them during their learning process. We concluded that students prefer to use known whatever messaging application, such as Telegram or WhatsApp, they are already using. They also expected chatbots to help them with their scheduling, to provide assignment grades and to search for class material. Only one of the applications mentioned, having them work with frequently asked questions, might actually require some understanding of natural language, although it can be formulated in such a way that is not needed.

% Maybe move it to a different place later
In the present work we aim to complement the aforementioned study with an analysis of teacher attitudes; teachers are the ones that actually would need to use, 
% Antonio - 'deploy' instead implement? Since in Computer Sciences implement and program are synonyms...
% JJ - deployment would be done by the university or whoever publishes the system. Deployment and implementation are two different things, anyway
% Antonio - Ok, but what I said is that 'implement' is a synonym of 'program' in this scope, and we write both in the same sentence as different things.
% And deploy does not make sense in this context. I will simply say use - JJ
configure and maybe program chatbots. Thus, we have performed a survey among 300 teachers asking about their attitudes towards the use of bots or chatbots in their classes. We analyse their responses, since we are interested in discussing implications and the impact for the development of future technology-enhanced tools and the Education institutions policies regarding the following research questions:
\begin{itemize}
\item{ RQ1 - Are teachers already using messaging apps in their classes? } %This was the second survey, not the  first survey. I wonder if maybe we could invert the presentation order - JJ
\item{ RQ2 - Which kind of chatbots would teachers find useful in their classes? }
\item{ RQ3 - Which kind of interaction do teachers prefer with their students? }
\item{ RQ4 - What kind of interaction media features do teachers value the most? }

@@

% Antonio - I include here the description of interaction features. Not sure if it is the best place for it. Move wherever it better fits.
Specifically, this survey asked the teachers, in addition to general information (such as gender, age, experience, etc), what are their feelings and preferences with regard to the use of communication tools with students, and even using bots or chatbots in the classroom.
In addition the survey asked about the most relevant \textit{Interaction media features} for the teachers with respect to the potential tool to use in class. These features are:
\begin{itemize}
\item \textit{Analytics}: the tool allows to extract data about students performance or development during the academic course on the teacher's subject.
\item \textit{Connectivity}: it is possible to link or connect the tool with other applications.
\item \textit{Familiarity}: it is something already known or previously used, so no need to learn something new.
\item \textit{Hidden Phone}: it is possible to use the tool without showing my phone number, normally using an alias or an own account independent from my physical telephone.
\item \textit{Horizontally}: all the actors are at the same level (for instance, teacher and students). There is not a hierarchy.
\item \textit{Official formation}: the university offers some training courses or talks about the tool (mainly for teachers). 
\item \textit{Pluggability}: the tool offers different pluggins to the teacher, so he/she can deploy his/her own implemented modules.
\item \textit{Sustainability}: teacher should not be worried about the tool maintenance.
\item \textit{Unidirectionality}: only the teacher can publish information in the channel, i.e. students are not allowed to ask or answer.
\item \textit{Officiality}: the tool has been granted by the university where the teacher works. This means normally that there is also technical support by university's staff.
\item \textit{Synchrony}: it is possible to publish/receive the data or information in several devices and all of them will contain or show the same.
\end{itemize}

@@

% Olia - TODO: check for repetitions in the following paragraph that you added. thank you.
The descriptive analysis in this case is done using a Chi-square test. For the different variables considered, some statistical significance was found for Gender, Sector, and academic Discipline. The values found with the test are: $p-value=0.0034$ for Gender, $p-value=0.0199$ for Sector, and $p-value=0.0211$ for Discipline.
The first column in Fig. \ref{figure:postcovid_chisq} shows the Pearson's residuals of the test for these three variables. Residuals show the difference between the observed and the expected values, with large residuals showing that variables are not truly independent. In this case, blue shows positive contribution and red negative contribution to the test, and saturation shows how large is this contribution in contrast to the expected value by chance. Tables on the right side show adjusted standardized residuals that according to \cite{agresti2013categorical}, if greater than +/-2 for cases with few cells, indicate lack of fit of $H_0$ (bolded in our table). Also, the table shows the percentage of contribution to the test of each cell.

@@

Answer to these questions are plotted in Figure \ref{figure:chatbots_types} grouped by gender, sector and experience. As it was a multiple choice question, the counts are over the total number of people that answered. One of the most interesting questions that can arise from the plots, and the should require further study, is: Are our professors really aware of the possibilities of chatbots in the classroom? The reason to ask ourselves this question is due to the fact of the small number of responses for other uses in comparison with the other approaches proposed. Another way to loook about it is that, unless you provide the specific instructions on how to use a tool, the teaching staff is already too busy or burn out to think about innovation.

@@

Since chatbots are intended to mediate or help in this interaction, it is essential to understand the kind of interaction teachers prefer. Chatbots should address those modes, and not others. In order to find out these modes, and thus answering RQ3, teachers were provided with a list of different kind of interactions that may take place among students and between students and the teacher, with the use of messaging apps. The findings are summarized in Table \ref{tab:kindinteractions}.

@@

 \item Keep it Simple. Media interaction requirements are simple and only a 2.4\% is requesting more features. This could be a sign of technological burn out produced by the previous and current courses where, due to the pandemic, the use of computer has increased significantly. 
\end{itemize}

@@

The key findings of this study shed the light on educators’ preference to use messaging applications that their institutions support.  Technology adoption literature often focus on users’ perceptions of the technology’s usefulness and ease of use as important prerequisites for successful adoption and utilization. Nevertheless, in higher education, institutions’ role in integrating these tools to their educational systems can improve the uptake of these applications and shape the social and educational experiences of their students.  This integration, to be fruitful, institutions have to ensure that these messaging applications not only are GDPR compliant to keep students’ data secure but also should provide IT support to all stakeholders who use the applications .

%This section is repeated
%\section*{Acknowledgements}

>>>>

Given the potential advantages that chatbots could bring to the classroom, as a previous step, we created two surveys for bachelor and master degree students at the University of Granada (Spain), which were answered by more than 250 students. The obtained results were analysed with the object of identifying students' actual needs, in \cite{MoraChatbots2021}. The survey asked about their preferences on tools/applications for chatting and messaging, how do they use them in the educational context, who do they like to be with in the class messaging groups, and what do they expect bots and chatbots to do to aid them during their learning process. We concluded that students prefer to use known whatever messaging application, such as Telegram or WhatsApp, they are already using. They also expected chatbots to help them with their scheduling, to provide assignment grades and to search for class material. Only one of the applications mentioned, having them work with frequently asked questions, might actually require some understanding of natural language, although it can be formulated in such a way that is not needed.
 In this sense, chatbot technology offers a great opportunity for the improvement of tutoring systems \cite{Daniel2016,agarwal2020}, as not all students are comfortable with face-to-face tutoring with the instructor. In fact, in many cases, some students experience stress because of their need to ask a question in front of the entire class. This is why many students refrain to silence in the classroom and tend to contact their teachers later via email. Consequently, this can lead to not only a delay in obtaining answers to their questions but also can cause a significant increase in their teachers’ workload, especially when factoring in the disproportion between the class size and the number of teachers. Hence, the chatbot technology has a potential to mitigate this problem by providing answers to students’ questions and facilitating a dynamic and autonomous learning experience \cite{Griol2014,kim2020}. In addition, using an automated system such as a chatbot, can draw teachers’ attention to topics that students struggle to comprehend or need further assistance in understanding. 
% Maybe move it to a different place later
In the present work we aim to complement the aforementioned study with an analysis of teacher attitudes; teachers are the ones that actually would need to use, 
% Olia-I believe we should avoid refereeing to “attitudes”. Also here, it is not mentioned that we explore their views towards the use of chat tools/ messaging apps. 
% Antonio - 'deploy' instead implement? Since in Computer Sciences implement and program are synonyms...
% JJ - deployment would be done by the university or whoever publishes the system. Deployment and implementation are two different things, anyway
% Antonio - Ok, but what I said is that 'implement' is a synonym of 'program' in this scope, and we write both in the same sentence as different things.
% And deploy does not make sense in this context. I will simply say use - JJ
configure and maybe program chatbots. Thus, we have performed a survey among 300 teachers asking about their attitudes towards the use of bots or chatbots in their classes. We analyse their responses, since we are interested in discussing implications and the impact for the development of future technology-enhanced tools and the Education institutions policies regarding the following research questions:
\begin{itemize}
\item{ RQ1 - Are teachers already using messaging apps in their classes? } %This was the second survey, not the  first survey. I wonder if maybe we could invert the presentation order - JJ
\item{ RQ2 - Which kind of chatbots would teachers find useful in their classes? }
\item{ RQ3 - Which kind of interaction do teachers prefer with their students? }
\item{ RQ4 - What kind of interaction media features do teachers value the most? }

@@

% Antonio - I include here the description of interaction features. Not sure if it is the best place for it. Move wherever it better fits.
% Olia- in the text below, we refer to "communucation tools", we should be consisten with the terms we use. Should we refere to communication tools? Messaging apps? Chat tools? 
Specifically, this survey asked the teachers, in addition to general information (such as gender, age, experience, etc), what are their feelings and preferences with regard to the use of communication tools with students, and even using bots or chatbots in the classroom.
In addition the survey asked about the most relevant \textit{Interaction media features} for the teachers with respect to the potential tool to use in class. These features are:
\begin{itemize}
\item \textit{Analytics}: the tool allows to extract data about students performance or development during the academic course on the teacher's subject.
\item \textit{Connectivity}: it is possible to link or connect the tool with other applications.
\item \textit{Familiarity}: it is something already known or previously used, so no need to learn something new.
\item \textit{Hidden Phone}: it is possible to use the tool without showing my phone number, normally using an alias or an own account independent from my physical telephone.
\item \textit{Horizontally}: all the actors are at the same level (for instance, teacher and students). There is not a hierarchy.
\item \textit{Official formation}: the university offers some training courses or talks about the tool (mainly for teachers). 
\item \textit{Pluggability}: the tool offers different pluggins to the teacher, so he/she can deploy his/her own implemented modules.
\item \textit{Sustainability}: teacher should not be worried about the tool maintenance.
\item \textit{Unidirectionality}: only the teacher can publish information in the channel, i.e. students are not allowed to ask or answer.
\item \textit{Officiality}: the tool has been granted by the university where the teacher works. This means normally that there is also technical support by university's staff.
\item \textit{Synchronous communication}: the interactions between teacher and students take place in real time, i.e. it is not a delayed communication scheme such as e-mail.
\end{itemize}

@@

% Olia - TODO: check for repetitions in the following paragraph that you added. thank you.
The first column in Fig. \ref{figure:postcovid_chisq} shows the Pearson's residuals of the test for these three variables. Residuals show the difference between the observed and the expected values, with large residuals showing that variables are not truly independent. In this case, blue shows positive contribution and red negative contribution to the test, and saturation shows how large is this contribution in contrast to the expected value by chance. Tables on the right side show adjusted standardized residuals that according to \cite{agresti2013categorical}, if greater than +/-2 for cases with few cells, indicate lack of fit of $H_0$ (bolded in our table). Also, the table shows the percentage of contribution to the test of each cell.

@@

Answer to these questions are plotted in Figure \ref{figure:chatbots_types} grouped by gender, sector and experience. As it was a multiple choice question, the counts are over the total number of people that answered. One of the most interesting questions that can arise from the plots, and the should require further study, is: Are teachers really aware of the possibilities of chatbots in the classroom? The reason to ask ourselves this question is due to the fact of the small number of responses for other uses in comparison with the other approaches proposed. Another way to look about it is that, unless you provide the specific instructions on how to use a tool, the teaching staff is already too busy or burn out to think about innovation.

@@

%Olia - move figure 7 elsewhere, not at the beginning of this section (RQ3). 
Since chatbots are intended to mediate or help in this interaction, it is essential to understand the kind of interaction teachers prefer. Chatbots should address those modes, and not others. In order to find out these modes, and thus answering RQ3, teachers were provided with a list of different kind of interactions that may take place among students and between students and the teacher, with the use of messaging apps. The findings are summarized in Table \ref{tab:kindinteractions}.

@@

 \item Keep it Simple. Media interaction requirements are simple and only a 2.4\% is requesting more features. This could be a sign of technological burn out produced by the previous and current courses where, due to the pandemic, the use of computer has increased significantly. 
\end{itemize}
%Olia: be consistent in the terms being used, we refer to teachers, then to professors. I would replace the word "professors" above with the word "teachers".

@@

The key findings of this study shed the light on educators’ preference to use messaging applications that their institutions support.  Technology adoption literature often focus on users’ perceptions of the technology’s usefulness and ease of use as important prerequisites for successful adoption and utilization. Nevertheless, in higher education, institutions’ role in integrating these tools to their educational systems can improve the uptake of these applications and shape the social and educational experiences of their students.  This integration, to be fruitful, institutions have to ensure that these messaging applications not only are GDPR compliant to keep students’ data secure but also should provide IT support to all stakeholders who use the applications .  
Moreover, this study clarifies the differences between the teachers’ perspectives and/or intentions and that of their students when using the messaging applications in higher education.  Teachers tend to adopt technologies supported by their institutions.  This could be to ensure that their universities can oversee their efforts in supporting their students during the learning process.  Another reason could be their familiarity with the technology provided by their institutions. On the other hand, students use the non-institution messaging applications to form informal discussion groups with their peers.  It is worth noting that peer support and collaboration is inseparable from learning \cite{ timmis2012constant } and correlates positively with higher retention in higher education \cite{ o2014front }.  Therefore, both perspectives are complementary and play different roles in promoting the learning process.
%This section is repeated
%\section*{Acknowledgements}
